The Ranked Nearest Neighbour Interchange space

### 2019-06-13 paper.tex
minor

%This lemma might be central  for induction proofs in RNNI space
%Is it correct and if it is, do we need to prove the statement, or is it rather obvious?
%The lemma implies another algorithm to compute distances (it would not give shortest paths explicitly, unlike alg:find_path)
\begin{lemma}
    Let $T_1$ and $T_2$ be two trees on $n+1$ taxa.
    Let $T_1|_n$ and $T_2|_n$ be these trees restricted to $n$ taxa, which means that the same taxon $n+1$ is deleted from both trees $T_1$ and $T_2$.
    Then it is $d(T_1|_n, T_2|_n) \leq d(T_1,T_2) - |r_{T_1}(p_{n+1}) - r_{T_2}(p_{n+1})|$ where $r_{T_i}(p_j)$ determines the rank of the parent of taxon $i$ in tree $T_i$.
\end{lemma}

@@

    Step($k \to k+1$):
    Let $T_1$ ad $T_2$ be trees on $n+1$ taxa with distance $d(T_1,T_2) = k$.
    Let $T_2'$ be the last tree on a shortest path from $T_1$ to $T_2$, i.e. $d(T_1,T_2') = k$.
    We can apply the induction hypothesis on $T_1$ and $T_2'$ and see that $d(T_1|_n, T_2'|_n) \leq d(T_1, T_2') - |r_{T_1}(p_{n+1}) - r_{T_2'}(p_{n+1})| := k - m$.
    Now we differ two cases:
    \begin{enumerate}
        \item $p_{n+1}$ changes its rank between $T_2$ and $T_2'$.

@@

        In this case it is $d(T_1|_n,T_2|_n) = d(T_1|_n, T_2'|_n) + 1 \leq (k-m)+ 1$.
        It also follows that $|r_{T_1}(p_{n+1}) - r_{T_2'}(p_{n+1})| = |r_{T_1}(p_{n+1}) - r_{T_2}(p_{n+1})|$ and therefore, we can conclude $d(T_1|_n,T_2|_n) \leq k-m+1 = k+1-m = d(T_1,T_2) -  |r_{T_1}(p_{n+1}) - r_{T_2}(p_{n+1})|$.
    \end{enumerate}
\end{proof}

>>>>

%This lemma might be central  for induction proofs in RNNI space
%Is it correct and if it is, do we need to prove the statement, or is it rather obvious?
%The lemma implies another algorithm to compute distances (it would not give shortest paths explicitly, unlike alg:find_path)
\begin{lemma}
    Let $T_1$ and $T_2$ be two trees on $n+1$ taxa.
    Let $T_1|_n$ and $T_2|_n$ be these trees restricted to $n$ taxa, i.e. the same taxon $n+1$ is deleted from both trees $T_1$ and $T_2$.
    Then it is $d(T_1|_n, T_2|_n) \leq d(T_1,T_2) - |r_{T_1}(p_{n+1}) - r_{T_2}(p_{n+1})|$ where $r_{T_i}(p_j)$ determines the rank of the parent of taxon $j$ in tree $T_i$.
\end{lemma}

@@

    Step($k \to k+1$):
    Let $T_1$ and $T_2$ be trees on $n+1$ taxa with distance $d(T_1,T_2) = k$.
    Let $T_2'$ be the last tree on a shortest path from $T_1$ to $T_2$, i.e. $d(T_1,T_2') = k$.
    We can apply the induction hypothesis on $T_1$ and $T_2'$ and see that $d(T_1|_n, T_2'|_n) \leq d(T_1, T_2') - |r_{T_1}(p_{n+1}) - r_{T_2'}(p_{n+1})| := k - m$.
    Now we differ two cases:
    \begin{enumerate}
        \item $p_{n+1}$ changes its rank between $T_2$ and $T_2'$.

@@

        In this case it is $d(T_1|_n,T_2|_n) \leq d(T_1|_n,T_2'|_n) + d(T_2'|_n, T_2|_n) \leq d(T_1|_n, T_2'|_n) + 1 \leq (k-m)+ 1$.
        It also follows that $|r_{T_1}(p_{n+1}) - r_{T_2'}(p_{n+1})| = |r_{T_1}(p_{n+1}) - r_{T_2}(p_{n+1})|$ and therefore, we can conclude $d(T_1|_n,T_2|_n) \leq k-m+1 = k+1-m = d(T_1,T_2) -  |r_{T_1}(p_{n+1}) - r_{T_2}(p_{n+1})|$.
    \end{enumerate}
\end{proof}
